---
title: 'Animal on the move: motion process concepts, biases and scales'
author: Alessandra Bertassoni; Cristiano H. G. M. Filho; Edgar Luiz de Lima; Felipe
  Pessoa da Silva; Filipe Guimarães Lima; Paulo De Marco Júnior
date: "25/11/2021"
output:
   html_document: default
---

## 1. Preparing the environment and data

Loading R packages.

```{r}
library(openxlsx)
library(tidyverse)
library(effectsize)
library(ggpubr)
```

Loading each reviewer's database.

```{r}
Ale<- read.xlsx("effort_Ale.xlsx")
cris<- read.xlsx("effort_Cristiano.xlsx")
ed<- read.xlsx("effort_Edgar.xlsx")
Feli<-read.xlsx("effort_Felipe.xlsx")
Fili<- read.xlsx("effort_Filipe.xlsx")
```

Union and standardization of databases.

```{r}
alldata<- rbind(Ale,cris, ed, Feli, Fili)

alldata$method<-alldata$method%>%str_replace_all( ";" , "/")%>%
  str_replace_all("_", " ")

alldata$metrica_summary<-alldata$metrica_summary%>%
  str_replace_all( ";" , "/")%>%str_replace_all( "," , "/")%>%
  str_replace_all("_", " ")

alldata$metrica_quali<-alldata$metrica_quali%>%
  str_replace_all( ";" , "/")%>%str_replace_all( "," , "/")%>%
  str_replace_all("_", " ")

View(alldata)# Viewing the data

write.xlsx(alldata, "alldata.xlsx", overwrite = T)# Save the data
```

After one more step of manual standardization with the presence of all reviewers, we arrived at the final database. Here we remove from the dataset the articles that were vetoed.

```{r}
## Loading final database
data<- read.xlsx("alldata_Final.xlsx")

## Removing vetoed articles
data<-data%>%replace(is.na(.), 0)%>%
  filter(method!= "vetted")%>%
  filter(method!=0 & obs!=0 )%>%
  select(PY,taxon_order1:disc_view)

## Viewing the data
View(data)
```

# 2. Biological scales x Study methods

Selecting columns to use in Biological Scales and study Methods Analysis.

```{r}
metlv<- data%>%select(theoretical,empirical_study,analytical_tools,
                       pop_biol, move_ecol, biogeog, land_ecol)
View(metlv)
odd<-as.data.frame(matrix(nrow = 8, ncol = 3))
resul<-as.data.frame(matrix(nrow = 8, ncol = 3))
```

Calculating the frequency tables of Biological Scales and study Methods Analysis.

```{r}
#Theoretical
ftpop<- ftable(metlv$theoretical, metlv$pop_biol)
ftpop<-ftpop[,c(2,1)]
ftmeco<- ftable(metlv$theoretical, metlv$move_ecol)
ftmeco<-ftmeco[,c(2,1)]
ftbigeo<- ftable(metlv$theoretical, metlv$biogeog)
ftbigeo<-ftbigeo[,c(2,1)]
ftlandeco<- ftable(metlv$theoretical, metlv$land_ecol)
ftlandeco<-ftlandeco[,c(2,1)]

```

Calculating the frequency tables of Biological Scales and study Methods Analysis.

```{r}
### Theoretical and Scales
ftpop<- ftable(metlv2$theoretical, metlv2$pop_biol)
ftpop<-ftpop[,c(2,1)]
ftmeco<- ftable(metlv$theoretical, metlv$move_ecol)
ftmeco<-ftmeco[,c(2,1)]
ftbigeo<- ftable(metlv$theoretical, metlv$biogeog)
ftbigeo<-ftbigeo[,c(2,1)]
ftlandeco<- ftable(metlv$theoretical, metlv$land_ecol)
ftlandeco<-ftlandeco[,c(2,1)]

### Odds Ratio
etpop<- oddsratio(ftpop)
odd[1,]<-etpop[,c(1,3,4)]
etmeco<- oddsratio(ftmeco)
odd[2,]<-etmeco[,c(1,3,4)]
etbigeo<-oddsratio(ftbigeo)
odd[3,]<-etbigeo[,c(1,3,4)]
etlandeco<- oddsratio(ftlandeco)
odd[4,]<-etlandeco[,c(1,3,4)]

### Empirical
fepop<- ftable(metlv$empirical_study, metlv$pop_biol)
fepop<-fepop[,c(2,1)]
femeco<- ftable(metlv$empirical_study, metlv$move_ecol)
femeco<-femeco[,c(2,1)]
febigeo<- ftable(metlv$empirical_study, metlv$biogeog)
febigeo<-febigeo[,c(2,1)]
felandeco<- ftable(metlv$empirical_study, metlv$land_ecol)
felandeco<-felandeco[,c(2,1)]

### Odds ratio
fepop<- oddsratio(fepop)
odd[5,]<-fepop[,c(1,3,4)]
femeco<- oddsratio(femeco)
odd[6,]<-femeco[,c(1,3,4)]
febigeo<-oddsratio(febigeo)
odd[7,]<-febigeo[,c(1,3,4)]
felandeco<- oddsratio(felandeco)
odd[8,]<-felandeco[,c(1,3,4)]


### Preparing the table.
View(odd)
colnames(odd)<- c("OR", "lw", "up")
sc<-c("Pop. Biology", "Movement Eco.", "Biogeography", "Land. Ecology")
tabela<-data.frame(sc,odd)
View(tabela)

write.xlsx(tabela, "BioScale.xlsx", overwrite = T)

tab<-tabela[1:4,]

```

Theoretical x Biological scales graphic.

```{r}
BS<-ggplot(tab, aes(x=sc,y= OR, ymin= lw, ymax= up))+ 
  geom_point(size=5)+
  geom_errorbar(data= tab, width= 0.05, size= 1)+
  coord_flip()+
  scale_y_continuous(breaks= seq(0.2,3.5, 0.3),limits = c(0.2,3.5))+
  geom_hline(yintercept = 1, linetype="dashed", size= 1)+
  xlab("Biological Scale")+
  ylab("Odds Ratio")+
  labs(tag = "a)", title = "Theoretical Study")+
  annotate("text", x = c(2.946), y =c(3.32), label = "*", size= 18 )+
  theme_classic() +
  theme(panel.grid.major = element_blank(),axis.text = element_text(colour = "Black",size=16), axis.text.y = element_text(colour = "Black",size=14),panel.grid.minor = element_blank(),axis.title = element_text(colour = "Black", size= 20), legend.title = element_blank(), legend.text = element_text(size = 18),plot.title = element_text(hjust = 0.5, size = 23),strip.background = element_blank(), strip.text = element_blank(),
        axis.line = element_line(size = 1),
        plot.tag = element_text(size = 16, face = "bold"))
```

# 3. Movement process and Study methods

Selecting columns to use in Movement process and Study Methods analysis.

```{r}
dfterms<- data%>%select(theoretical,empirical_study,dispersial_bin,dispersion_bin,
  displacement_bin,movement_bin, migration_bin)
View(dfterms)



odd2<-as.data.frame(matrix(nrow = 10, ncol = 3))
resul2<-as.data.frame(matrix(nrow = 10, ncol = 3))
```

Calculating the frequency tables of Motion process and Study Methods.

```{r}
ftable(dfterms$theoretical, dfterms$empirical_study)
dfterms<- dfterms%>%mutate(excl= theoretical+empirical_study)%>%filter(excl!=2)
View(dfterms)

#Theoretical
ftdsal<- ftable(dfterms$theoretical, dfterms$dispersial_bin)
ftdsal<- ftdsal[,c(2,1)]
ftdsion<- ftable(dfterms$theoretical, dfterms$dispersion_bin)
ftdsion<- ftdsion[,c(2,1)]
ftdispla<- ftable(dfterms$theoretical, dfterms$displacement_bin)
ftdispla<- ftdispla[,c(2,1)]
ftmov<- ftable(dfterms$theoretical, dfterms$movement_bin)
ftmov<- ftmov[,c(2,1)]
ftmigra<- ftable(dfterms$theoretical, dfterms$migration_bin)
ftmigra<- ftmigra[,c(2,1)]

### Effect size estimate.
etsal<- oddsratio(ftdsal)
odd2[1,]<-etsal[,c(1,3,4)]
etsion<- oddsratio(ftdsion)
odd2[2,]<-etsion[,c(1,3,4)]
etispla<- oddsratio(ftdispla)
odd2[3,]<-etispla[,c(1,3,4)]
etmov<- oddsratio(ftmov)
odd2[4,]<-etmov[,c(1,3,4)]
etmigra<- oddsratio(ftmigra)
odd2[5,]<-etmigra[,c(1,3,4)]

##Empirical
fedsal<- ftable(dfterms$empirical_study, dfterms$dispersial_bin)
fedsal<- fedsal[,c(2,1)]
fedsion<- ftable(dfterms$empirical_study, dfterms$dispersion_bin)
fedsion<- fedsion[,c(2,1)]
fedispla<- ftable(dfterms$empirical_study, dfterms$displacement_bin)
fedispla<- fedispla[,c(2,1)]
femov<- ftable(dfterms$empirical_study, dfterms$movement_bin)
femov<- femov[,c(2,1)]
femigra<- ftable(dfterms$empirical_study, dfterms$migration_bin)
femigra<- femigra[,c(2,1)]

## Odds ratio
eesal<- oddsratio(fedsal)
odd2[6,]<-eesal[,c(1,3,4)]
eesion<- oddsratio(fedsion)
odd2[7,]<-eesion[,c(1,3,4)]
eeispla<- oddsratio(fedispla)
odd2[8,]<-eeispla[,c(1,3,4)]
eemov<- oddsratio(femov)
odd2[9,]<-eemov[,c(1,3,4)]
eemigra<- oddsratio(femigra)
odd2[10,]<-eemigra[,c(1,3,4)]

### Preparing the table
View(odd2)
colnames(odd2)<- c("OR", "lw", "up")
sc2<-c("Dispersal", "Dispersion", "Displacement", "Movement", "Migration")
tabela2<-data.frame(sc2,odd2)
View(tabela2)

write.xlsx(tabela2, "MotionProcess.xlsx", overwrite = T)
```

Theoretical x Motion process graphic.

```{r}
MP<-ggplot(tabela2, aes(x=sc2,y= OR, ymin= lw, ymax= up))+ 
  geom_point(size=5)+
  geom_errorbar(data= tabela2, width= 0.05, size= 1)+
  coord_flip()+
  scale_y_continuous(breaks= seq(0.0,5.0, 0.5),limits = c(0.0,5.0))+
  geom_hline(yintercept = 1, linetype="dashed", size= 1)+
  xlab("Motion process")+
  ylab("Odds Ratio")+
  labs(tag = "b)", title = "Theoretical Study")+
  annotate("text", x = c(0.9), y =c(0.8), label = "*", size= 18 )+
  #annotate("text", x = c(7), y =c(1),label = "*", size= 9 )+
  theme_classic() +
  theme(panel.grid.major = element_blank(),axis.text = element_text(colour = "Black",size=16), axis.text.y = element_text(colour = "Black",size=14),panel.grid.minor = element_blank(),axis.title = element_text(colour = "Black", size= 20), legend.title = element_blank(), legend.text = element_text(size = 18),plot.title = element_text(hjust = 0.5, size = 23),strip.background = element_blank(), strip.text = element_blank(),
        axis.line = element_line(size = 1),
        plot.tag = element_text(size = 16, face = "bold"))
```

Uniting the graphics.

```{r}
ggarrange(BS,MP, nrow = 1,ncol = 2)

ggsave(filename = "Figure_5.pdf", width = 55, height = 22, units = "cm", dpi = 600)
```

# 4. Taxon analysis of Biological scale x Motion Process

Excluding studies that are both theoretical and empirical for mammals.

```{r}
ftable(metlv$theoretical, metlv$empirical_study)
metlv2<- metlv%>%mutate(excl= theoretical+empirical_study)%>%filter(excl!=2)
View(metlv2)
```

Estimating the frequency table.

```{r}
#Terms
dispersial<- ftable(dados$taxon_group, dados$dispersial_bin)
dispersial<- dispersial[c(2,1),]
dispersion<- ftable(dados$taxon_group, dados$dispersion_bin)
dispersion<- dispersion[c(2,1),]
displacement<- ftable(dados$taxon_group, dados$displacement_bin)
displacement<- displacement[c(2,1),]
movement<- ftable(dados$taxon_group, dados$movement_bin)
movement<- movement[c(2,1),]
migration<- ftable(dados$taxon_group, dados$migration_bin)
migration<- migration[c(2,1),]

#Escala
biol<- ftable(dados$taxon_group, dados$pop_biol)
biol<- biol[c(2,1),]
movecol<- ftable(dados$taxon_group, dados$move_ecol)
movecol<- movecol[c(2,1),]
biogeo<- ftable(dados$taxon_group, dados$biogeog)
biogeo<- biogeo[c(2,1),]
land<- ftable(dados$taxon_group, dados$land_ecol)
land<- land[c(2,1),]

odd3<-as.data.frame(matrix(nrow = 9, ncol = 3))
resul3<-as.data.frame(matrix(nrow = 9, ncol = 3))

```

Odds ratio estimate for motion process and biological scales.

```{r}
#Motion process
orsal<- oddsratio(dispersial)
odd3[1,]<-orsal[,c(1,3,4)]
orsion<- oddsratio(dispersion)
odd3[2,]<-orsion[,c(1,3,4)]
ordispla<- oddsratio(displacement)
odd3[3,]<-ordispla[,c(1,3,4)]
ormov<- oddsratio(movement)
odd3[4,]<-ormov[,c(1,3,4)]
ormigra<- oddsratio(migration)
odd3[5,]<-ormigra[,c(1,3,4)]

#Biological scale
orbiol<- oddsratio(pbiol)
odd3[6,]<-orbiol[,c(1,3,4)]
ormov<- oddsratio(pmov)
odd3[7,]<-ormov[,c(1,3,4)]
orbiogeo<- oddsratio(pbiogeo)
odd3[8,]<-orbiogeo[,c(1,3,4)]
orland<- oddsratio(pland)
odd3[9,]<-orland[,c(1,3,4)]
```

Preparing the table.

```{r}
ts2<- c("Terms","Terms", "Terms","Terms","Terms", "Scale","Scale","Scale","Scale")
ter2<- c("Dispersal", "Dispersion", "Displacement", "Movement", "Migration",
        "Pop. Biology", "Movement Eco.", "Biogeography", "Land. Ecology" )

resul4<-data.frame(ts2, ter2, odd3)
colnames(resul4)<-  c("TS", "TheSca", "OR", "LW", "UP")
resul4[2,5]<- 0.0
View(resul4)

bs<-filter(resul4,TS=="Scale")
mp<-filter(resul4,TS=="Terms")
write.xlsx(resul4,"Taxon.xlsx", overwrite = T)
```

Biological scale x Theoretical study for mammals graphic.

```{r}
BS2<- ggplot(bs, aes(x=TheSca,y= OR, ymin= LW, ymax= UP))+ 
  geom_point(size=5)+
  geom_errorbar(data= bs, width= 0.05, size= 1)+
  coord_flip()+
  scale_y_continuous(breaks= seq(0.0,4.0, 0.5),limits = c(0.0,4.0))+
  geom_hline(yintercept = 1, linetype="dashed", size= 1)+
  xlab("Biological Scale")+
  ylab("Odds Ratio")+
  labs(tag= "a)" ,title = "Mammals")+
  annotate("text", x = c(2.9), y =c(0.8), label = "*", size= 18 )+
  theme_classic() +
  theme(panel.grid.major = element_blank(),axis.text = element_text(colour = "Black",size=16), axis.text.y = element_text(colour = "Black",size=14),panel.grid.minor = element_blank(),axis.title = element_text(colour = "Black", size= 20), legend.title = element_blank(), legend.text = element_text(size = 18),plot.title = element_text(hjust = 0.5, size = 23),strip.background = element_blank(), axis.line = element_line(size = 1),
        plot.tag = element_text(size = 16, face = "bold"))

```

Motion process x Theoretical study for mammals graphic.

```{r}
MP2<- ggplot(mp, aes(x=TheSca,y= OR, ymin= LW, ymax= UP))+ 
  geom_point(size=5)+
  geom_errorbar(data= mp, width= 0.05, size= 1)+
  coord_flip()+
  scale_y_continuous(breaks= seq(0.0,7.5, 0.5),limits = c(0.0,7.5))+
  geom_hline(yintercept = 1, linetype="dashed", size= 1)+
  xlab("Motion process")+
  ylab("Odds Ratio")+
  labs(tag= "b)" ,title = "Mammals")+
  annotate("text", x = c(0.9), y =c(7.5), label = "*", size= 18 )+
  annotate("text", x = c(1.9), y =c(0.2), label = "*", size= 18 )+
  theme_classic() +
  theme(panel.grid.major = element_blank(),axis.text = element_text(colour = "Black",size=16), axis.text.y = element_text(colour = "Black",size=14),panel.grid.minor = element_blank(),axis.title = element_text(colour = "Black", size= 20), legend.title = element_blank(), legend.text = element_text(size = 18),plot.title = element_text(hjust = 0.5, size = 23),strip.background = element_blank(), axis.line = element_line(size = 1),
        plot.tag = element_text(size = 16, face = "bold"))

```

```{r}
ggarrange(BS2,MP2, nrow = 1,ncol = 2)

ggsave(filename = "Figure_6.pdf", width = 60, height = 22, units = "cm", dpi = 600)
```
